%!TEX root = ../main.tex
%********************************
\chapter{Results and Demonstration}
\section{Outputs}
\subsection{Speech-to-Text and Intent Recognition}
The application successfully processes voice commands into text using Whisper and Faster-Whisper. The transcription maintains high accuracy for both Finnish and English inputs, demonstrating multilingual support. Intent recognition using Mistral Instruct accurately interprets commands such as:
\begin{itemize}
    \item \textbf{“Create an image of a sunset.”} → Generates a new image using SDXL-Lightning.
    \item \textbf{“Edit the image to add a mountain.”} → Modifies the existing image with Instruct-Pix2Pix.
    \item \textbf{“Animate this image into a video.”} → Generates a video sequence using Stable Video Diffusion.
\end{itemize}

\subsubsection{Screenshots}
\begin{itemize}
    \item Show step-by-step outputs (speech-to-text, intent recognition logs, and structured JSON commands).
    \item Highlight intermediate previews for edits and final outputs for videos.
\end{itemize}

\subsection{Generated Images}
\begin{itemize}
    \item Examples of high-resolution images created with SDXL-Lightning, demonstrating:
    \begin{itemize}
        \item Sharpness and quality even with short prompts.
        \item Ability to handle complex scenes with multiple objects or stylistic variations.
    \end{itemize}
    \item Visuals illustrating edits made using Instruct-Pix2Pix, such as:
    \begin{itemize}
        \item Adding elements (e.g., “add a tree”).
        \item Changing colors or moods (e.g., “make the image look like a sunset”).
    \end{itemize}
\end{itemize}

\subsection{Video Outputs}
\begin{itemize}
    \item Short video clips created with Stable Video Diffusion, showing smooth animations and transitions from static inputs.
\end{itemize}

\subsection{Demonstration Use Cases}
\begin{itemize}
    \item \textbf{Scenario 1:} Generating and editing a landscape image, followed by creating an animated version.
    \item \textbf{Scenario 2:} Creating a poster design, modifying text and colors, and adding animation effects.
\end{itemize}

\section{Performance Observations}
\subsection{Latency Measurements}
\begin{itemize}
    \item \textbf{Speech-to-Text:}
    \begin{itemize}
        \item \textbf{Whisper (Faster-Whisper):} Achieves transcription times averaging ~1-2 seconds for short audio inputs (~5-10 seconds long).
        \item Longer inputs (~30 seconds) complete in ~6-8 seconds with acceptable accuracy.
    \end{itemize}
    \item \textbf{Intent Recognition:}
    \begin{itemize}
        \item \textbf{Mistral Instruct (Q4 Quantization)} processes commands in ~200–300 milliseconds, balancing speed and resource usage.
    \end{itemize}
    \item \textbf{Image Generation:}
    \begin{itemize}
        \item \textbf{SDXL-Lightning} generates 1024x1024 images in ~1.5–3 seconds, depending on prompt complexity.
        \item Edits using \textbf{Instruct-Pix2Pix} complete in ~2-4 seconds, including rendering updates.
    \end{itemize}
    \item \textbf{Video Generation:}
    \begin{itemize}
        \item \textbf{Stable Video Diffusion} produces 2–4 second videos in approximately 10–15 seconds, depending on the frame count and resolution.
    \end{itemize}
\end{itemize}

\subsection{GPU Usage and Resource Observations}
\begin{itemize}
    \item \textbf{Memory Consumption:}
    \begin{itemize}
        \item Whisper and Mistral (quantized) comfortably fit within 8–12 GB VRAM, even on mid-range GPUs.
        \item SDXL-Lightning and Stable Video Diffusion require 16–24 GB VRAM for optimal performance.
        \item Dynamic model unloading successfully reduces memory pressure, allowing models to run sequentially on systems with limited VRAM.
    \end{itemize}
    \item \textbf{Resource Bottlenecks:}
    \begin{itemize}
        \item \textbf{Multiple Users:} Concurrent requests cause latency to increase significantly due to model reloading delays and GPU memory constraints.
        \item \textbf{Batch Processing:} Larger batches (e.g., generating multiple outputs at once) slow down performance and may cause out-of-memory errors without careful optimization.
    \end{itemize}
\end{itemize}

\subsection{Observations on Accuracy and Quality}
\begin{itemize}
    \item \textbf{Speech Recognition:} Whisper maintains high transcription accuracy, especially for shorter inputs. However, background noise or overlapping speech can slightly reduce quality, suggesting room for noise filtering improvements in future updates.
    \item \textbf{Image Outputs:}
    \begin{itemize}
        \item \textbf{SDXL-Lightning} produces sharp and detailed images, particularly for abstract or artistic styles.
        \item Realistic outputs may require prompt tuning to balance detail and coherence.
    \end{itemize}
    \item \textbf{Video Outputs:} Stable Video Diffusion generates smooth transitions, but longer animations may lose coherence without additional conditioning inputs.
\end{itemize}